---
title: "hw3-8105"
author: "Chenshuo Pan"
output: github_document
---

```{r}
library(tidyverse)
library(ggplot2)
library(patchwork)
```



# Question1:

```{r}
#read dataset
library(p8105.datasets)
data("instacart")
```

**write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations.**

This dataset has `r nrow(instacart)` rows and `r ncol(instacart)` columns ,The key variables include:`r names(instacart)`.

(From the description of this data on the course website,we can describe the key variables as follows)

`order_id`: order identifier

`product_id`: product identifier

`add_to_cart_order`: order in which each product was added to cart

`reordered`: 1 if this prodcut has been ordered by this user in the past, 0 otherwise

`user_id`: customer identifier

`eval_set`: which evaluation set this order belongs in (Note that the data for use in this class is exclusively from the “train” eval_set)

`order_number`: the order sequence number for this user (1=first, n=nth)

`order_dow`: the day of the week on which the order was placed

`order_hour_of_day`: the hour of the day on which the order was placed

`days_since_prior_order`: days since the last order, capped at 30, NA if order_number=1

`product_name`: name of the product

`aisle_id`: aisle identifier

`department_id`: department identifier

`aisle`: the name of the aisle

`department`: the name of the department


```{r}
#an example of observation data

instacart%>%filter(user_id ==21)
```

Take user No. 21 as an example. This user purchased 6 products, 5 of which this customer has purchased in the past. The order number is 34, and the order date is 12 o'clock on Monday. It has been 28 days since this customer last placed an order. This data also includes information such as the type of product purchased by the customer, the department to which the product belongs, and the aisle.



**How many aisles are there, and which aisles are the most items ordered from?**

```{r}
aisle_rank <-instacart%>%
  group_by(aisle_id)%>%
  summarise(aisle_number=n())%>%
  arrange(desc(aisle_number))

head(aisle_rank,1)
```


There are totally `r nrow(aisle_rank)` aisles, among those `aisle_id = 83` which is `r instacart%>%filter(aisle_id == 83)%>%head(1)%>%select(aisle) ` are the most items. The number is `r aisle_rank%>%select(aisle_number)%>%head(1)` 



**Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.**

```{r}
#keep aisles with more than 10000 items

instacart_10000 <- aisle_rank%>%
  filter(aisle_number >= 10000)

  
instacart_10000%>%
  mutate(aisle_id = reorder(aisle_id,desc(aisle_number)))%>%
  ggplot(aes(x = aisle_id,y = aisle_number))+
  geom_bar(stat = "identity")+
  labs(title = "Number of items in each aisle",
       x = "Aisle ID",
       y = "Number of items in this aisle")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```




**Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.**

```{r}
most3_item <- instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits"))%>%
  group_by(aisle,product_name)%>%
  summarise(product_number = n(),.groups = 'keep')%>%
  arrange(aisle,desc(product_number))%>%
  group_by(aisle)%>%
  top_n(3,product_number)


most3_item
```



**Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).**

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream"))%>%
  group_by(product_name,order_dow)%>%
  summarise(mean_hour = mean(order_hour_of_day),.groups = 'keep')%>%
  mutate(order_dow = case_when(
    order_dow == 0 ~ "Sunday",
    order_dow == 1 ~ "Monday",
    order_dow == 2 ~ "Tuesday",
    order_dow == 3 ~ "Wednesday",
    order_dow == 4 ~ "Thursday",
    order_dow == 5 ~ "Friday",
    order_dow == 6 ~ "Saturday",
  ))%>%
  pivot_wider(names_from = "order_dow",values_from = mean_hour)
```




# Question2:


```{r}
data("brfss_smart2010")
```


**do some data cleaning:**
**format the data to use appropriate variable names;**
**focus on the “Overall Health” topic**
**include only responses from “Excellent” to “Poor”**
**organize responses as a factor taking levels ordered from “Poor” to “Excellent”**

```{r}
brfss_clean <- brfss_smart2010%>%
  janitor::clean_names()%>%
  rename(state = locationabbr,state_county = locationdesc) %>%
  filter(topic == "Overall Health") %>%
  filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor"))%>%
  mutate(response = factor(response,
         levels = c("Excellent", "Very good", "Good", "Fair", "Poor")))
```




**In 2002, which states were observed at 7 or more locations? What about in 2010?**

```{r}
brfss_2002 <- brfss_clean %>%
  filter(year == 2002)%>%
  group_by(state)%>%
  summarise(n=length(unique(state_county)))%>%
  filter(n >= 7)

brfss_2002

brfss_2010 <- brfss_clean %>%
  filter(year == 2010)%>%
  group_by(state)%>%
  summarise(n=length(unique(state_county)))%>%
  filter(n >= 7)

brfss_2010
```


There are totally `r nrow(brfss_2002)` states were observed at 7 or more locations in 2002 .They are `r pull(brfss_2002,state)`

There are totally `r nrow(brfss_2010)` states were observed at 7 or more locations in 2010 .They are `r pull(brfss_2010,state)`










**Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).**

```{r}
state_response <- brfss_clean%>%
  filter(response == "Excellent")%>%
  group_by(year,state,state_county)%>%
  summarise(average_value = mean(data_value),.groups = "keep")

state_response%>%
  group_by(year,state)%>%
  summarise(year_value = mean(average_value),.groups = "keep")%>%
  ggplot(aes(x = year,y = year_value,color = state,group = state))+
  geom_line()+
  labs(title = "spaghetti plot of average value over time within a state",
       x = "Year",
       y = "Yearly Data Value") 
  #facet_grid(~state)

```

Comment:


**Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.**


```{r}
violin_2006<- brfss_clean%>%
  filter(year==2006 | state == "NY") %>%
  ggplot(aes(x = response, y = data_value))+
  geom_violin(aes(fill = response))+
  viridis::scale_fill_viridis(discrete = TRUE)

violin_2010<- brfss_clean%>%
  filter(year==2010 | state == "NY") %>%
  ggplot(aes(x = response, y = data_value))+
  geom_violin(aes(fill = response))+
  viridis::scale_fill_viridis(discrete = TRUE)

violin_2006 / violin_2010 
```

Comment:


# Question3:


```{r}
covar <- read.csv("./data/nhanes_covar.csv",skip = 4,header = TRUE)
accel <- read.csv("./data/nhanes_accel.csv",header = TRUE)
```


**Load, tidy, merge, and otherwise organize the data sets. Your final dataset should include all originally observed variables; exclude participants less than 21 years of age, and those with missing demographic data; and encode data with reasonable variable classes (i.e. not numeric, and using factors with the ordering of tables and plots in mind).**

```{r}
covar_clean <- covar%>%
  janitor::clean_names()%>%
  mutate(sex = case_when(
    sex == 1 ~ "male",
    sex == 2 ~ "female"
  ),
  education = case_when(
    education == 1 ~"Less than high school",
    education == 2 ~"High school equivalent",
    education == 3 ~"More than high school"
  )) %>%
  filter(age >=21)%>%
  na.omit()



accel_clean <- accel%>%
  janitor::clean_names()

```

```{r}
combine_table<-covar_clean%>%
  inner_join(accel_clean,by = "seqn")%>%
  pivot_longer(min1:min1440,names_to = "time", values_to = "mims")%>%
  na.omit()
```

**Produce a reader-friendly table for the number of men and women in each education category, and create a visualization of the age distributions for men and women in each education category. Comment on these items.**
```{r}
combine_table%>%
  group_by(education,sex)%>%
  summarise(n=n())



#age value only in covar
covar_clean%>%
  ggplot(aes(x = age , fill = sex,))+
  geom_density(alpha = 0.5)+
  facet_grid(~education)+ 
  viridis::scale_fill_viridis(discrete = TRUE)
  
```
Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences. Comment on your plot.


```{r}
combine_table%>%
  group_by(sex,age,education,seqn)%>%
  summarise(total = sum(mims))%>%
  ggplot(aes(x = age , y = total,color = sex))+
  geom_point(alpha = 0.5)+
  geom_smooth(se = FALSE)+
  facet_grid(.~education)
```





Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences.


```{r}
as.data.frame(combine_table%>%
  filter(education == "Less than high school")%>%
  group_by(sex,time)%>%
  summarise(mean_min = mean(mims),.groups = 'keep')%>%
  apply(2,function(time) gsub("min", "", time)))%>%
  mutate(time = as.numeric(time))%>%
  arrange(time)%>%
  ggplot(aes(x = time, y = mean_min,color = sex ))+
  geom_point()+
  scale_y_discrete(breaks = c(0, 20, 40, 60,80,100))+
  theme_minimal()+
  labs(title = "Number of items in each aisle",
       x = "Aisle ID",
       y = "Number of items in this aisle")
```



```{r}

```












